Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web

### 2024-11-25 autovi_paper.qmd
reword VSS description

---
title: "Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web"
format:
  quarto-anzjs-pdf:
    keep-tex: true  
    shorttitle: "Automated Residual Plot Assessment"
    double-space: true
    line-numbers: true
    # You can specify extra preamble in the file below.
    include-in-header:
      - header.tex
date: today
date-format: "YYYY" # used to define volumeyear
author:
  - name: Weihao Li
    email: weihao.li@monash.edu
    corresponding: true
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0003-4959-106X
  - name: Dianne Cook
    email: dicook@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0002-3813-7155
  - name: Emi Tanaka
    email: emi.tanaka@anu.edu.au
    affiliations: 
      - name: The Australian National University
        department: Biological Data Science Institute
        address: 46 Sullivan's Creek Road, ACT 2600, Australia
    orcid: 0000-0002-1455-259X
  - name: Susan VanderPlas
    email: emi.tanaka@anu.edu.au
    affiliations:
      - name: University of Nebraska
        department: Department of Statistics
        address: Hardin Hall, 3310 Holdrege St Suite 340, Lincoln, NE 68583, United States
    orcid: 0000-0002-3803-0972
  - name: Klaus Ackermann
    email: Klaus.Ackermann@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0001-7693-8538
# Note that the Journal requires that a paper must begin with a
# "Summary" not an "Abstract".  This is automatically taken care
# of by the anzsauth document style.  So even though the following
# says "abstract" the heading "Summary" will appear in
# the processed version.
abstract: |
  Visually assessing residual plots is a common advice for linear model diagnostics, however this approach requires manual human evaluation and thereby is not scalable for assessing many models. Human evaluation also suffer from the potential for inconsistent decisions from different analysts. Using a lineup protocol, where the residual plot is embedded among null plots, can help to alleviate inconsistency, but requires even more human effort. This is the type of task that in today's world might employ a robot to do the tedious work for a human. Here we describe a new R package that includes a computer vision model for automated assessment of residual plots, and an accompanying Shiny app for ease of use. For a user-provided sample of residuals, it predicts a measure of visual signal strength (VSS) and provides a suite of supporting information to assist the analyst decide on the appropriateness their model fit. 
# Note that "keywords" should not include words and phrases
# that form part of the title of the paper.
# Also keywords (with the exception of proper names and
# certain abbreviations that conventionally appear all in
# capital letters) should *not* be capitalised.
keywords: [initial data analysis, statistical graphics, data visualization, visual inference, computer vision, machine learning, hypothesis testing, regression analysis, model diagnostics]
bibliography: bibliography.bib  
#nocite: |
---

@@

**Visual signal strength** (VSS) refers to how prominently a specific set of visual patterns appears in an image. In the context of regression model diagnostics, it describes the clarity of visual patterns on a diagnostic plot that may indicate model violations. This concept can be categorized as weak, moderate, or strong. However, in our study, we treat it as a continuous positive real variable. Importantly, its interpretation depends on how it is linked to a function of the data or the underlying data generating process. Consequently, the meaning of visual signal strength varies across different model classes and even within the same model, depending on the chosen function.

In @li2024automated, visual signal strength was defined as an estimate of a distance measure that quantifies the disparity between the residual distribution of a fitted classical normal linear regression model and a reference distribution. This measure is based on the Kullback-Leibler (KL) divergence:

@@

Here, $p(.)$ and $q(.)$ are the probability density functions of the reference residual distribution $P$ and the true residual distribution $Q$, respectively.

This distance measure requires knowledge of the true residuals' data generating process, which is typically unknown. Therefore, it must be estimated. To address this, a computer vision model was trained to perform the estimation.

>>>>

---
title: "Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web"
format:
  quarto-anzjs-pdf:
    keep-tex: true  
    shorttitle: "Automated Residual Plot Assessment"
    double-space: true
    line-numbers: true
    # You can specify extra preamble in the file below.
    include-in-header:
      - header.tex
date: today
date-format: "YYYY" # used to define volumeyear
author:
  - name: Weihao Li
    email: weihao.li@monash.edu
    corresponding: true
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0003-4959-106X
  - name: Dianne Cook
    email: dicook@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0002-3813-7155
  - name: Emi Tanaka
    email: emi.tanaka@anu.edu.au
    affiliations: 
      - name: The Australian National University
        department: Biological Data Science Institute
        address: 46 Sullivan's Creek Road, ACT 2600, Australia
    orcid: 0000-0002-1455-259X
  - name: Susan VanderPlas
    email: emi.tanaka@anu.edu.au
    affiliations:
      - name: University of Nebraska
        department: Department of Statistics
        address: Hardin Hall, 3310 Holdrege St Suite 340, Lincoln, NE 68583, United States
    orcid: 0000-0002-3803-0972
  - name: Klaus Ackermann
    email: Klaus.Ackermann@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0001-7693-8538
# Note that the Journal requires that a paper must begin with a
# "Summary" not an "Abstract".  This is automatically taken care
# of by the anzsauth document style.  So even though the following
# says "abstract" the heading "Summary" will appear in
# the processed version.
abstract: |
  Visually assessing residual plots is a common advice for linear model diagnostics, however this approach requires manual human evaluation and thereby is not scalable for assessing many models. Human evaluation also suffer from the potential for inconsistent decisions from different analysts. Using a lineup protocol, where the residual plot is embedded among null plots, can help to alleviate inconsistency, but requires even more human effort. This is the type of task that in today's world might employ a robot to do the tedious work for a human. Here we describe a new R package that includes a computer vision model for automated assessment of residual plots, and an accompanying Shiny app for ease of use. For a user-provided sample of residuals, it predicts a measure of visual signal strength and provides a suite of supporting information to assist the analyst decide on the appropriateness their model fit. 
# Note that "keywords" should not include words and phrases
# that form part of the title of the paper.
# Also keywords (with the exception of proper names and
# certain abbreviations that conventionally appear all in
# capital letters) should *not* be capitalised.
keywords: [initial data analysis, statistical graphics, data visualization, visual inference, computer vision, machine learning, hypothesis testing, regression analysis, model diagnostics]
bibliography: bibliography.bib  
#nocite: |
---

@@

Visual signal strength (VSS) refers to how prominently a specific set of visual patterns appears in an image. In the context of regression model diagnostics, it describes the clarity of visual patterns on a diagnostic plot that may indicate model violations. This concept can be categorized as weak, moderate, or strong. However, in our study, we treat it as a continuous positive real variable. Importantly, its interpretation depends on how it is linked to a function of the data or the underlying data generating process. Consequently, the meaning of visual signal strength varies across different model classes and even within the same model, depending on the chosen function.

In @li2024automated, VSS was defined as an estimate of a distance measure that quantifies the disparity between the residual distribution of a fitted classical normal linear regression model and a reference distribution. This measure is based on the Kullback-Leibler (KL) divergence:

@@

here, $p(.)$ and $q(.)$ are the probability density functions of the reference residual distribution $P$ and the true residual distribution $Q$, respectively.

This distance measure requires knowledge of the true residual distribution, which is typically unknown. Therefore, a computer vision model was used for estimating the residual distribution [see @li2024automated for more details].

### 2024-11-25 autovi_paper.tex
reword VSS description

\usepackage{pmboxdraw}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{anzsj}
\usepackage{bookmark}

@@

\begin{abstract}
Visually assessing residual plots is a common advice for linear model
diagnostics, however this approach requires manual human evaluation and
thereby is not scalable for assessing many models. Human evaluation also
suffer from the potential for inconsistent decisions from different
analysts. Using a lineup protocol, where the residual plot is embedded
among null plots, can help to alleviate inconsistency, but requires even
more human effort. This is the type of task that in today's world might
employ a robot to do the tedious work for a human. Here we describe a
new R package that includes a computer vision model for automated
assessment of residual plots, and an accompanying Shiny app for ease of
use. For a user-provided sample of residuals, it predicts a measure of
visual signal strength (VSS) and provides a suite of supporting
information to assist the analyst decide on the appropriateness their
model fit.
\end{abstract}

@@

\textbf{Visual signal strength} (VSS) refers to how prominently a
specific set of visual patterns appears in an image. In the context of
regression model diagnostics, it describes the clarity of visual
patterns on a diagnostic plot that may indicate model violations. This
concept can be categorized as weak, moderate, or strong. However, in our
study, we treat it as a continuous positive real variable. Importantly,
its interpretation depends on how it is linked to a function of the data
or the underlying data generating process. Consequently, the meaning of
visual signal strength varies across different model classes and even
within the same model, depending on the chosen function.

In \citet{li2024automated}, visual signal strength was defined as an
estimate of a distance measure that quantifies the disparity between the
residual distribution of a fitted classical normal linear regression
model and a reference distribution. This measure is based on the
Kullback-Leibler (KL) divergence:

@@

Here, \(p(.)\) and \(q(.)\) are the probability density functions of the
reference residual distribution \(P\) and the true residual distribution
\(Q\), respectively.

This distance measure requires knowledge of the true residuals' data
generating process, which is typically unknown. Therefore, it must be
estimated. To address this, a computer vision model was trained to
perform the estimation.

>>>>

\usepackage{pmboxdraw}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{anzsj}
\usepackage{bookmark}

@@

\begin{abstract}
Visually assessing residual plots is a common advice for linear model
diagnostics, however this approach requires manual human evaluation and
thereby is not scalable for assessing many models. Human evaluation also
suffer from the potential for inconsistent decisions from different
analysts. Using a lineup protocol, where the residual plot is embedded
among null plots, can help to alleviate inconsistency, but requires even
more human effort. This is the type of task that in today's world might
employ a robot to do the tedious work for a human. Here we describe a
new R package that includes a computer vision model for automated
assessment of residual plots, and an accompanying Shiny app for ease of
use. For a user-provided sample of residuals, it predicts a measure of
visual signal strength and provides a suite of supporting information to
assist the analyst decide on the appropriateness their model fit.
\end{abstract}

@@

Visual signal strength (VSS) refers to how prominently a specific set of
visual patterns appears in an image. In the context of regression model
diagnostics, it describes the clarity of visual patterns on a diagnostic
plot that may indicate model violations. This concept can be categorized
as weak, moderate, or strong. However, in our study, we treat it as a
continuous positive real variable. Importantly, its interpretation
depends on how it is linked to a function of the data or the underlying
data generating process. Consequently, the meaning of visual signal
strength varies across different model classes and even within the same
model, depending on the chosen function.

In \citet{li2024automated}, VSS was defined as an estimate of a distance
measure that quantifies the disparity between the residual distribution
of a fitted classical normal linear regression model and a reference
distribution. This measure is based on the Kullback-Leibler (KL)
divergence:

@@

here, \(p(.)\) and \(q(.)\) are the probability density functions of the
reference residual distribution \(P\) and the true residual distribution
\(Q\), respectively.

This distance measure requires knowledge of the true residual
distribution, which is typically unknown. Therefore, a computer vision
model was used for estimating the residual distribution \citep[see][ for
more details]{li2024automated}.
